---
title: "Project"
author: "Daniel Walker"
date: "4/30/2019"
output: html_document
---
```{r echo=TRUE, message=FALSE}
# always clean up R environment
rm (list = ls ())

#install.packages('RCurl')

# load all packages here
library (mdsr)
library (dplyr) # for data manipluation
library (ggplot2) # data visualization
library (reshape2) # reshaping 

# user-defined functions here (if any)

# load data

```

```{r}
# read data https://www.kaggle.com/starbucks/starbucks-menu
drinks_expanded <- read.csv('starbucks_drinkMenu_expanded.csv')
food <- read.csv('starbucks-menu-nutrition-food.csv')

# as the data seems to contains some trademarks and accent, we will first remove them
drinks_expanded$Beverage_category <- iconv (drinks_expanded$Beverage_category, "latin1", "ASCII", sub = "byte")
drinks_expanded$Beverage <- iconv (drinks_expanded$Beverage, "latin1", "ASCII", sub = "byte")
drinks_expanded$Beverage_prep <- iconv (drinks_expanded$Beverage_prep, "latin1", "ASCII", sub = "byte")

drinks_expanded$Beverage_category <- gsub ('(<[a-z0-9]+>)','',drinks_expanded$Beverage_category)
drinks_expanded$Beverage <- gsub ('(<[a-z0-9]+>)','',drinks_expanded$Beverage)
drinks_expanded$Beverage_prep <- gsub ('(<[a-z0-9]+>)','',drinks_expanded$Beverage_prep)
```

### Quick view of Data
```{r}
head (food)
head (drinks_expanded)
```

### Plot Strar Bucks Food (Calories and Fat)
```{r}
# red line shows recomended meal amount of Calories for food
# blue line shows MAX recomended daily amount of Fat for Food

# https://www.uccs.edu/healthcircle/sites/healthcircle/files/inline-files/Meal%20Planning.pdf
# https://my.clevelandclinic.org/health/articles/11208-fat-what-you-need-to-know

plot (Calories~Fat, 
     ylab = 'Calories in Starbucks Food', 
     xlab = 'Fat in Starbucks Food',
     main = 'Caleries vs Fat for Starbucks Food', 
     data = food)
abline (v=22, col='blue')
abline (h=500, col='red')

food %>%
  select (Food.name,Calories,Fat) %>%
  filter (Calories>=500,Fat>=22) %>% 
  count (Food.name) %>% 
  ungroup () %>% 
  arrange (desc (n))

# Unhealthiest Option
food %>%
  select (Food.name, Calories, Fat) %>%
  filter (Calories>=600, Fat>=28) %>% 
  count (Food.name) %>% 
  ungroup () %>% 
  arrange (desc (n))

if (FALSE){
"
Green Goddess Avocado Salad			
Lentils & Vegetable Protein Bowl with Brown Rice		
Roasted Turkey & Dill Havarti Sandwich
"
}
```

### Plot Strar Bucks Drink (Calories and Total Fat)
```{r}
# red line shows recomended meal amount of Calories for food

# https://www.uccs.edu/healthcircle/sites/healthcircle/files/inline-files/Meal%20Planning.pdf
# https://my.clevelandclinic.org/health/articles/11208-fat-what-you-need-to-know

plot (Calories~Total_Fat, 
     ylab = 'Calories in Starbucks Drinks', 
     xlab = 'Total Fat in Starbucks Food',
     main = 'Caleries vs Total Fat for Starbucks Drinks', 
     data = drinks_expanded)
abline (h = 500, col = 'red')
```

### Plot Strar Bucks Drink (Sugars and Total Fat)
```{r}
# red line shows MAX recomended meal amount of sugar for food

# http://sugarscience.ucsf.edu/the-growing-concern-of-overconsumption.html

plot (Sugars ~ Total_Fat, 
     xlab = 'Fat in Starbucks Drinks', 
     ylab = 'Total Sugar in Starbucks Food',
     main = 'Total Fat Vs Sugar for Starbucks Drinks', 
     data = drinks_expanded)
abline (h = 38, col = 'red')
```

### Plot Strar Bucks Drink (Sugars and Caffine)
```{r, warning=FALSE, message=FALSE}
# Blue line Max recomendation of sugar
# Red line Max recomendation of caffine

# https://www.fda.gov/consumers/consumer-updates/spilling-beans-how-much-caffeine-too-much

sudo <- drinks_expanded %>%
  mutate (clean_caff = as.numeric (as.character (drinks_expanded$Caffeine)))
sudo[is.na (sudo)] <- 0

plot (clean_caff~Sugars, 
     ylab = 'Caffeine in Starbucks Drinks', 
     xlab = 'Sugar in Starbucks Food',
     main = 'Total Fat Vs Sugar for Starbucks Drinks', 
     data = sudo)
abline (h=400, col='red')
abline (v=38, col='blue')

# Venti Coffee
sudo %>%
  select (Beverage_category, clean_caff) %>%
  filter (clean_caff >= 400)

sudo %>%
  select (Beverage_category, Sugars) %>%
  filter (Sugars >= 38) %>% 
  count (Beverage_category) %>% 
  ungroup () %>% 
  arrange (desc (n))

sudo_sugar <- sudo %>%
  select (Beverage_category, Sugars) %>%
  filter (Sugars >= 38) 

# 37.60331% of drinks have higher than rated sugar
length (sudo_sugar[,1]) / length (drinks_expanded$Beverage_category)

if (FALSE){
"
Worst drinks ordered by violation of sugar contents
Frappuccino Blended Coffee				
Signature Espresso Drinks				
Tazo Tea Drinks  
"
}

```

#### Checking for missing values
It doesnt seems the data contains any missing values. However, Caffeine variable in the `Drinks Expanded` dataset seems to be a mixed of numeric and character. As it can be seen, there are values with value `varies` or `Varies` in the data. Removed them from the dataset and will format it into numeric variable.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
summary (drinks_expanded)
summary (food)
drinks_expanded <- subset (drinks_expanded, !Caffeine %in% c ('Varies', 'varies', ''))
drinks_expanded$Caffeine <- as.numeric (as.character (drinks_expanded$Caffeine))
```

#### Distribution of the variables in each of the datasets
Drinks expanded dataset
```{r ggplot, fig.width=14, fig.height=8, warning=FALSE, message=FALSE, echo=FALSE}
drinks_expanded %>% 
    select (-contains ("Beverage")) %>% 
    mutate (dummy = 1) %>% 
    melt (id = 'dummy') %>%
    mutate (value=as.numeric (value)) %>% 
    ggplot (aes (value)) + 
    geom_histogram (color = 'white', fill = 'red') + 
    facet_wrap (~variable, scales = 'free', ncol = 5) + 
    theme_bw () +
    labs (x = 'Value',y = 'Frequency',title = 'Drinks Expanded Variables Histograms')+
    theme(
        legend.text = element_text(size=11),
        legend.position = "top",
        legend.title = element_blank(),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        axis.title.x = element_text(size=12),
        axis.title.y = element_text(size=12),
        strip.text.y = element_text(size=12),
        strip.text.x = element_text(size=12),
        plot.title = element_text(size=15))
```

###Food dataset
```{r fig.width=14, fig.height=4, warning=FALSE, message=FALSE, echo=FALSE}
food %>% 
    select (-contains("Food")) %>% 
    mutate (dummy=1) %>% 
    melt (id='dummy') %>%
    mutate (value = as.numeric(value)) %>% 
    ggplot (aes (value)) + 
    geom_histogram (color = 'white', fill = 'red') + 
    facet_wrap (~variable, scales = 'free', ncol = 5) + 
    theme_bw () +
    labs(x = 'Value', y = 'Frequency',title = 'Food Variables Histograms')+
    theme(
        legend.text = element_text(size=11),
        legend.position = "top",
        legend.title = element_blank(),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        axis.title.x = element_text(size=12),
        axis.title.y = element_text(size=12),
        strip.text.y = element_text(size=12),
        strip.text.x = element_text(size=12),
        plot.title=element_text(size=15))
```

### Supervised learning
Fit linear model to predict number of Calories in a drink based on other variables such as
Total_Fat, Trans_Fat. Sugars etc.
5 folds cross validation Used and will use `MAPE` (mean average error percentage) as the perforamce metrics. Once we have the `MAPE` for each fold, I will simply take the average of all folds and this will be the final `MAPE` 
```{r warning=FALSE, message=FALSE}
# randomly shuffle the data, set seed to produce the same resultss
# in addition, remove the first 3 columns which contains the Beverage descriptions
set.seed (380)
drinks_expanded <- drinks_expanded[sample(nrow(drinks_expanded)),] %>% 
  select (-contains("Beverage"))

# create 5 equally size folds
K <- 5
folds <- cut (seq (1, nrow (drinks_expanded)), breaks = K, labels = FALSE)

# Perform 5 folds cross validation
MAPE <- c ()
for (i in 1:K){
    index <- which (folds==i, arr.ind=TRUE)
    train <- drinks_expanded[-index, ]
    test <- drinks_expanded[index, ]
    model <- lm (Calories ~., data=train)
    prediction <- predict (model,test)
    MAPE[i] <- abs (test$Calories / prediction - 1)
}

paste0 ('MAPE for 5 folds: ', round (mean (MAPE)*100, 2),'%')

```

### Unspurviesd learning
for the unspurviesd part, I will use the food dataset.
We will use "Hierarchical Clustering" in order to find clusters of similiar foods based on
`Calories`, `Fat`, `Carb`, `Fiver` and `Protein`
```{r fig.width=24, fig.height=15, warning=FALSE, message=FALSE}
# remove the food column 
df <- food %>% 
  select (-Food.name)
# set food column as the row names
rownames (df) <- food$Food.name

# scale the data and create distance metrics using euclidean distance.
dd <- dist (scale (df), method = "euclidean")

# fit hierarchical clustering using "ward.D2" method
hc <- hclust (dd, method = "ward.D2")

# plot the results
par (cex = 1, mar = c (2, 2, 4, 1))
plot (hc, 
     hang = -1, 
     cex = 1.4, 
     xlab = "", 
     sub = "", 
     axes = FALSE, 
     ylab = '', 
     main = 'Food HierarChial Cluster Diagram',
     cex.main = 2)
```

### Simulation
```{r}
# https://www.accessdata.fda.gov/scripts/interactivenutritionfactslabel/carbohydrate.html

set.seed (380)
starbucks_order <- function (food_size = 1, drink_size = 1){
  food_samp <- sample (food ,food_size, replace=TRUE)
  drink_samp <- sample (drinks_expanded, drink_size, replace=TRUE)
  
  total_sugar <- sum (drink_samp$Sugars)
  total_caffine <- sum (drink_samp$Caffeine)
  total_calories <- sum (food_samp$Calories) + sum (drink_samp$Calories)
  total_fat <- sum (food_samp$Fat) + sum (drink_samp$Total_Fat)
  total_carbo <- sum (food_samp$Carb) + sum (drink_samp$Total_Carbohydrates)
  
  # index 1 sugar
  # index 2 calories
  # index 3 fat
  # index 4 carbo
  # index 5 caffine
  return (c (total_sugar, total_calories, total_fat, total_carbo, total_caffine))
}

food_orders <- 3
drink_orders <- 3
n <- 365
final_sim <- do (n) * starbucks_order (food_orders,drink_orders)
head (final_sim)

hist (final_sim[,1],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Sugar Intake at Starbucks',
     main = 'Daily Sugar Amount (g)')
abline (v = 38, col = 'red')

hist (final_sim[,2],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Calories Intake at Starbucks',
     main = 'Daily Calories Amount')

hist (final_sim[,3],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Fat Intake at Starbucks',
     main = 'Daily Fat Amount (g)')
abline (v = 65, col = 'red')

hist (final_sim[,4],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Carbohydrates Intake at Starbucks',
     main = 'Daily Carbohydrates Amount (g)')
abline (v=300,col='red')

hist (final_sim[,5],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Caffine Intake at Starbucks',
     main = 'Daily Caffine Amount (mg)')
abline (v=400, col='red')

```


