---
title: "STAT 380 Final Project"
subtitle: "Spring 2019"
author: "Sanchit Goel and Daniel Walker"
date: "Due April 30, 2019"
output: html_notebook
---

# Front matter
```{r}
# always clean up R environment
rm(list = ls())

# load all packages here
library (dplyr)
library (tidyr)
library (ggplot2)
library (caret)
library (readr)
library (tidyverse)
library (ggthemes)
library (stringr)
library (repr)
library (plotly)
library (corrplot)
library (caTools)
library (rpart)
library (rpart.plot)
library (mdsr)
library (reshape2)

# Setting a random seed
set.seed(380)
```

```{r echo = TRUE}
# User Defined Function:
replace_drop_na <- function(arg_1) {
   arg_1[arg_1 == '-'] <- NA
   print(paste0("Pre-Cleaning Records: ", nrow(arg_1)))

  # Dropping NA values records
  result <- arg_1 %>%
  drop_na()

  print(paste0("Cleaned Records: ", nrow(result)))
  return(result)
}
```


# Loading and Cleaning the Datasets
### McDonalds Dataset
```{r}
mcd <- read.csv(file = "nutrition_facts_mcdonalds_menu.csv", header = TRUE, sep = ",")
mcd <- replace_drop_na(mcd)

# Subsetting drinks from the menu
drinks.oz <- mcd[str_detect(mcd$Serving.Size, " fl oz.*"), ]
drinks.ml <- mcd[str_detect(mcd$Serving.Size, 'carton'), ]

# Converting ounces to mililiters and further rounding the values
drinks.oz$Serving.Size <- 
  round(as.numeric(gsub(" fl oz.*", "", drinks.oz$Serving.Size)) * 29.5735, 0)
drinks.ml$Serving.Size <- 
  round(as.numeric(gsub(".*\\((.*)\\ ml).*", "\\1", drinks.ml$Serving.Size)), 0)

# Keeping the values and removing characters
food.g <- mcd[str_detect(mcd$Serving.Size, 'g'), ] 
food.g$Serving.Size <- 
  round(as.numeric(gsub(".*\\((.*)\\ g).*", "\\1", food.g$Serving.Size)), 0)

# Combining drinks and food menu
# While adding a new feature for drink or food
cleaned_mcd <- rbind(drinks.oz, drinks.ml)
cleaned_mcd$Type <- rep("Drinks", nrow(cleaned_mcd))
food.g$Type <- rep("Food", nrow(food.g))
cleaned_mcd <- rbind(cleaned_mcd, food.g)

keeps <- c("Item", "Calories", "Total.Fat", "Carbohydrates", "Dietary.Fiber", "Protein", "Type")
cleaned_mcd <- cleaned_mcd[keeps]
colnames(cleaned_mcd)[colnames(cleaned_mcd)=="Total.Fat"] <- "Total_Fat"
colnames(cleaned_mcd)[colnames(cleaned_mcd)=="Dietary.Fiber"] <- "Fiber"

cleaned_mcd$Calories <- as.numeric(as.character(cleaned_mcd$Calories))
cleaned_mcd$Total_Fat <- as.numeric(as.character(cleaned_mcd$Total_Fat))
cleaned_mcd$Carbohydrates <- as.numeric(as.character(cleaned_mcd$Carbohydrates))
cleaned_mcd$Fiber <- as.numeric(as.character(cleaned_mcd$Fiber))
cleaned_mcd$Protein <- as.numeric(as.character(cleaned_mcd$Protein))
```

### Starbucks Dataset
```{r}
# Food Dataset
sbux_f <- read.csv(file = "nutrition_facts_starbucks_food_menu.csv", header = TRUE, sep = ",")
sbux_f <- replace_drop_na(sbux_f)

colnames(sbux_f)[colnames(sbux_f) == "ï..Food.Name"] <- "Item"
colnames(sbux_f)[colnames(sbux_f) == "Fat..g."] <- "Total_Fat"
colnames(sbux_f)[colnames(sbux_f) == "Carb...g."] <- "Carbohydrates"
colnames(sbux_f)[colnames(sbux_f) == "Fiber..g."] <- "Fiber"
colnames(sbux_f)[colnames(sbux_f) == "Protein..g."] <- "Protein"

# Drink Dataset
sbux_d <- read.csv(file = "nutrition_facts_starbucks_drink_menu.csv", header = TRUE, sep = ",")
sbux_d <- replace_drop_na(sbux_d)

sbux_d <- sbux_d %>%
  select(-one_of(c('Sodium')))
colnames(sbux_d)[colnames(sbux_d) == "X"] <- "Item"
colnames(sbux_d)[colnames(sbux_d) == "Fat..g."] <- "Total_Fat"
colnames(sbux_d)[colnames(sbux_d) == "Carb...g."] <- "Carbohydrates"
colnames(sbux_d)[colnames(sbux_d) == "Fiber..g."] <- "Fiber"

# Adding a new column named Type
sbux_d$Type <- rep("Drinks", nrow(sbux_d))
sbux_f$Type <- rep("Food", nrow(sbux_f))
cleaned_sbux <- rbind(sbux_f, sbux_d)

cleaned_sbux$Calories <- as.numeric(as.character(cleaned_sbux$Calories))
cleaned_sbux$Total_Fat <- as.numeric(as.character(cleaned_sbux$Total_Fat))
cleaned_sbux$Carbohydrates <- as.numeric(as.character(cleaned_sbux$Carbohydrates))
cleaned_sbux$Fiber <- as.numeric(as.character(cleaned_sbux$Fiber))
cleaned_sbux$Protein <- as.numeric(as.character(cleaned_sbux$Protein))
```

### Starbucks Drinks Expanded Dataset
```{r}
drinks_expanded <- read.csv('starbucks_drink_menu_expanded.csv', header = TRUE, sep = ",")

# as the data seems to contains some trademarks and accent, we will first remove them
drinks_expanded$Beverage_category <- iconv(drinks_expanded$Beverage_category, "latin1", "ASCII", sub = "byte")
drinks_expanded$Beverage <- iconv(drinks_expanded$Beverage, "latin1", "ASCII", sub = "byte")
drinks_expanded$Beverage_prep <- iconv(drinks_expanded$Beverage_prep, "latin1", "ASCII", sub = "byte")

drinks_expanded$Beverage_category <- gsub('(<[a-z0-9]+>)', '', drinks_expanded$Beverage_category)
drinks_expanded$Beverage <- gsub('(<[a-z0-9]+>)', '', drinks_expanded$Beverage)
drinks_expanded$Beverage_prep <- gsub('(<[a-z0-9]+>)', '', drinks_expanded$Beverage_prep)

colnames(drinks_expanded)[colnames(drinks_expanded) =="Total.Fat..g."] <- "Total_Fat"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Trans.Fat..g."] <- "Trasn_Fat"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Saturated.Fat..g."] <- "Saturated_Fat"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Sodium..mg."] <- "Sodium"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Total.Carbohydrates..g."] <- "Total_Carbohydrates"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Cholesterol..mg."] <- "Cholesterol"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Dietary.Fibre..g."] <- "Dietary_Fibre"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Sugars..g."] <- "Sugars"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Protein..g."] <- "Protein"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Vitamin.A....DV."] <- "Vitamin_A"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Vitamin.C....DV."] <- "Vitamin_C"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Calcium....DV."] <- "Calcium"
colnames(drinks_expanded)[colnames(drinks_expanded) =="Iron....DV."] <- "Iron"
colnames(drinks_expanded)[colnames(drinks_expanded) == "Caffeine..mg."] <- "Caffeine"

drinks_expanded$Vitamin_A <- gsub( "%", "", as.character(drinks_expanded$Vitamin_A))
drinks_expanded$Vitamin_A <- as.numeric(as.character(drinks_expanded$Vitamin_A))

drinks_expanded$Vitamin_C <- gsub( "%", "", as.character(drinks_expanded$Vitamin_C))
drinks_expanded$Vitamin_C <- as.numeric(as.character(drinks_expanded$Vitamin_C))

drinks_expanded$Calcium <- gsub( "%", "", as.character(drinks_expanded$Calcium))
drinks_expanded$Calcium <- as.numeric(as.character(drinks_expanded$Calcium))

drinks_expanded$Iron <- gsub( "%", "", as.character(drinks_expanded$Iron))
drinks_expanded$Iron <- as.numeric(as.character(drinks_expanded$Iron))

drinks_expanded$Caffeine <- gsub( "Varies", "", as.character(drinks_expanded$Caffeine))
drinks_expanded$Caffeine <- gsub( "varies", "", as.character(drinks_expanded$Caffeine))
drinks_expanded$Caffeine <- as.numeric(as.character(drinks_expanded$Caffeine))
drinks_expanded$Total_Fat <- as.numeric(as.character(drinks_expanded$Total_Fat))

drinks_expanded <- replace_drop_na(drinks_expanded)
```

### Joining the Mcdonalds and Starbucks Dataset
```{r}
# Joining the Food and Drink Menu for Starbucks and Mcdonalds
cleaned_sbux$Brand <- rep("Starbucks", nrow(cleaned_sbux))
cleaned_mcd$Brand <- rep("McDonalds", nrow(cleaned_mcd))
joined_data <- rbind(cleaned_mcd, cleaned_sbux)
```

### McDonalds Data Summary
```{r echo = TRUE}
summary(cleaned_mcd)
```

### Starbucks Data Summary
```{r echo = TRUE}
summary(cleaned_sbux)
```

### Starbucks Drinks Expanded Summary
```{r}
summary(drinks_expanded)
```

### Joined Data Summary
```{r echo = TRUE}
summary(joined_data)
```

```{r}
# Freeing up Memory of redundant data
rm('mcd', 'sbux_d', 'keeps', 'drinks.ml', 'drinks.oz', 'food.g')
```


## McDonalds:
#### Calories Histogram
```{r echo = TRUE}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(x = Calories)) +
  geom_histogram(aes(y = ..density..), fill = "lightblue", binwidth = 40, color = "darkgray") + 
  scale_x_continuous(breaks = seq(min(cleaned_mcd$Calories), max(cleaned_mcd$Calories), by = 200)) + 
  geom_density()
```

#### Calories BoxPlot
```{r echo = TRUE}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(x = reorder(Type, Calories), y = Calories)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = NULL)
```

#### Correlation Plot
```{r echo = TRUE}
cr <- cor(cleaned_mcd[, 2:6])
corrplot(cr, method = 'color')
rm('cr')
```

#### Calories Vs Total Fat Scatter Plot
```{r echo = TRUE}
plot(cleaned_mcd$Total_Fat, cleaned_mcd$Calories, main = "Scatter Plot", xlab = "Total Fat", ylab = "Calories")
abline(lm(cleaned_mcd$Calories ~ cleaned_mcd$Total_Fat))
```

#### Colries vs Protein Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(y = Calories, x = Protein)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(y = Calories, x = Carbohydrates)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(y = Calories, x = Total_Fat)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

### Model Training
#### Splitting the dataset
```{r}
split <- sample.split(cleaned_mcd$Calories, SplitRatio = 0.80)
train <- subset(cleaned_mcd, split == "TRUE")
test <- subset(cleaned_mcd, split == "FALSE")
```

#### Linear Regression Model
```{r}
mcd_lm_model <- lm(Calories ~ Total_Fat + Protein + Carbohydrates, data = train)
summary(mcd_lm_model)
```

#### Plots to explain the model
```{r echo = TRUE}
plot(mcd_lm_model, which = 1:4)
```

#### Predictions vs Actual
```{r echo = TRUE}
predictions <- predict(mcd_lm_model, test)
plot(test$Calories, type = "l", lty = 1.8, col = "blue")
lines(predictions, type = "l", col = "red")
```

#### Future Predictions
```{r}
predict(mcd_lm_model, data.frame(Total_Fat = 20, Protein = 18, Carbohydrates = 33))
```

#### Decision Tree
```{r}
mcd_dt_model <- rpart(Type ~ Calories + Total_Fat + Protein + Carbohydrates, data = train)
prp(mcd_dt_model, type = 1, extra = 3, main = "Decision Tree")

predictions <- data.frame(data = factor(predict(mcd_dt_model, test, type = "class")), type = "prediction")
real <- data.frame(data = factor(test$Type), type = "real")
conf_data <- rbind(predictions, real)

# Check if the levels are identical
identical(levels(conf_data[conf_data$type == "prediction", 1]), levels(conf_data[conf_data$type == "real", 1]))

confusionMatrix(conf_data[conf_data$type == "prediction", 1], conf_data[conf_data$type == "real",1], dnn = c("Prediction", "Reference"))
```

```{r}
# Freeing up Memory of redundant data
rm('predictions', 'real', 'conf_data', 'split', 'test', 'train')
```

## Starbucks:
#### Calories Histogram
```{r echo = TRUE}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(x = Calories)) +
  geom_histogram(aes(y = ..density..), fill = "lightblue", binwidth = 40, color = "darkgray") + 
  scale_x_continuous(breaks = seq(min(cleaned_sbux$Calories), max(cleaned_sbux$Calories), by = 200)) + 
  geom_density()
```

#### Calories BoxPlot
```{r echo = TRUE}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(x = reorder(Type, Calories), y = Calories)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = NULL)
```

#### Correlation Plot
```{r echo = TRUE}
cr <- cor(cleaned_sbux[, 2:6])
corrplot(cr, method = 'color')
rm('cr')
```

#### Calories Vs Total Fat Scatter Plot
```{r echo = TRUE}
plot(cleaned_sbux$Total_Fat, cleaned_sbux$Calories, main = "Scatter Plot", xlab = "Total Fat", ylab = "Calories")
abline(lm(cleaned_sbux$Calories ~ cleaned_sbux$Total_Fat))
```

#### Colries vs Protein Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(y = Calories, x = Protein)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(y = Calories, x = Carbohydrates)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(y = Calories, x = Total_Fat)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

### Model Training
#### Splitting the dataset
```{r}
split <- sample.split(cleaned_sbux$Calories, SplitRatio = 0.80)
train <- subset(cleaned_sbux, split == "TRUE")
test <- subset(cleaned_sbux, split == "FALSE")
```

#### Linear Regression Model
```{r}
sbux_lm_model <- lm(Calories ~ Total_Fat + Protein + Carbohydrates, data = train)
summary(sbux_lm_model)
```

#### Plots to explain the model
```{r echo = TRUE}
plot(sbux_lm_model, which = 1:4)
```

#### Predictions vs Actual
```{r echo = TRUE}
predictions <- predict(sbux_lm_model, test)
plot(test$Calories, type = "l", lty = 1.8, col = "blue")
lines(predictions, type = "l", col = "red")
```

#### Future Predictions
```{r}
predict(sbux_lm_model, data.frame(Total_Fat = 20, Protein = 18, Carbohydrates = 33))
```

#### Decision Tree
```{r}
sbux_dt_model <- rpart(Type ~ Calories + Total_Fat + Protein + Carbohydrates, data = train)
prp(sbux_dt_model, type = 1, extra = 3, main = "Decision Tree")

predictions <- data.frame(data = factor(predict(sbux_dt_model, test, type="class")), type = "prediction")
real <- data.frame(data = factor(test$Type), type = "real")
conf_data <- rbind(predictions, real)

# Check if the levels are identical
identical(levels(conf_data[conf_data$type == "prediction", 1]), levels(conf_data[conf_data$type == "real", 1]))

confusionMatrix(conf_data[conf_data$type == "prediction", 1], conf_data[conf_data$type == "real",1], dnn = c("Prediction", "Reference"))
```

```{r}
# Freeing up Memory of redundant data
rm('predictions', 'real', 'conf_data', 'split', 'test', 'train')
```

## Joined Data:
#### Calories Histogram
```{r echo = TRUE}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(joined_data, aes(x = Calories)) +
  geom_histogram(aes(y = ..density..), fill = "lightblue", binwidth = 40, color = "darkgray") + 
  scale_x_continuous(breaks = seq(min(joined_data$Calories), max(joined_data$Calories), by = 200)) + 
  geom_density()
```

#### Calories BoxPlot
```{r echo = TRUE}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(joined_data, aes(x = reorder(Brand, Calories), y = Calories)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = NULL)
```

#### Correlation Plot
```{r echo = TRUE}
cr <- cor(joined_data[, 2:6])
corrplot(cr, method = 'color')
rm('cr')
```

#### Calories Vs Total Fat Scatter Plot
```{r echo = TRUE}
plot(joined_data$Total_Fat, joined_data$Calories, main = "Scatter Plot", xlab = "Total Fat", ylab = "Calories")
abline(lm(joined_data$Calories ~ joined_data$Total_Fat))
```

#### Colries vs Protein Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(joined_data, aes(y = Calories, x = Protein)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Brand, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(joined_data, aes(y = Calories, x = Carbohydrates)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Brand, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = TRUE}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(joined_data, aes(y = Calories, x = Total_Fat)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Brand, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

### Model Training
#### Splitting the dataset
```{r}
split <- sample.split(joined_data$Calories, SplitRatio = 0.80)
train <- subset(joined_data, split == "TRUE")
test <- subset(joined_data, split == "FALSE")
```

#### Linear Regression Model
```{r}
lm_model <- lm(Calories ~ Total_Fat + Protein + Carbohydrates, data = train)
summary(lm_model)
```

#### Plots to explain the model
```{r echo = TRUE}
plot(lm_model, which = 1:4)
```

#### Predictions vs Actual
```{r echo = TRUE}
predictions <- predict(lm_model, test)
plot(test$Calories, type = "l", lty = 1.8, col = "blue")
lines(predictions, type = "l", col = "red")
```

#### Future Predictions
```{r}
predict(lm_model, data.frame(Total_Fat = 20, Protein = 18, Carbohydrates = 33))
```

#### Decision Tree
```{r}
dt_model <- rpart(Type ~ Calories + Total_Fat + Protein + Carbohydrates, data = train)
prp(dt_model, type = 1, extra = 3, main = "Decision Tree")

predictions <- data.frame(data = factor(predict(dt_model, test, type="class")), type = "prediction")
real <- data.frame(data = factor(test$Type), type = "real")
conf_data <- rbind(predictions, real)

# Check if the levels are identical
identical(levels(conf_data[conf_data$type == "prediction", 1]), levels(conf_data[conf_data$type == "real", 1]))

confusionMatrix(conf_data[conf_data$type == "prediction", 1], conf_data[conf_data$type == "real",1], dnn = c("Prediction", "Reference"))
```

```{r}
# Freeing up Memory of redundant data
rm('predictions', 'real', 'conf_data', 'split', 'test', 'train')
```

### Plot StarBucks Food (Calories and Fat)
```{r}
# Red line shows recomended meal amount of Calories for food
# Blue line shows MAX recomended daily amount of Fat for Food

# https://www.uccs.edu/healthcircle/sites/healthcircle/files/inline-files/Meal%20Planning.pdf
# https://my.clevelandclinic.org/health/articles/11208-fat-what-you-need-to-know

plot(Calories~Total_Fat, 
     ylab = 'Calories in Starbucks Food', 
     xlab = 'Fat Food in Starbucks Food',
     main = 'Caleries vs Fat for Starbucks Food', 
     data = sbux_f)
abline(v = 22, col = 'blue')
abline(h = 500, col = 'red')

sbux_f %>%
  select(Item, Calories, Total_Fat) %>%
  filter(Calories >= 500, Total_Fat >= 22) %>% 
  count(Item) %>% 
  ungroup() %>% 
  arrange(desc(n))

# Unhealthiest Option
sbux_f %>%
  select(Item, Calories, Total_Fat) %>%
  filter(Calories >= 600, Total_Fat >= 28) %>% 
  count(Item) %>% 
  ungroup() %>% 
  arrange(desc(n))

# Green Goddess Avocado Salad     
# Lentils & Vegetable Protein Bowl with Brown Rice    
# Roasted Turkey & Dill Havarti Sandwich
```

### Plot StarBucks Drink (Calories and Total Fat)
```{r}
# Red line shows recomended meal amount of Calories for food

# https://www.uccs.edu/healthcircle/sites/healthcircle/files/inline-files/Meal%20Planning.pdf
# https://my.clevelandclinic.org/health/articles/11208-fat-what-you-need-to-know

plot(Calories~Total_Fat, 
     ylab = 'Calories in Starbucks Drinks', 
     xlab = 'Total Food in Starbucks Food',
     main = 'Caleries vs Total Fat for Starbucks Drinks', 
     data = drinks_expanded)
abline(h = 500, col = 'red')
```


### Plot StarBucks Drink (Sugars and Total Fat)
```{r}
# Red line shows MAX recomended meal amount of sugar for food

# http://sugarscience.ucsf.edu/the-growing-concern-of-overconsumption.html

plot(Sugars~Total_Fat, 
     xlab = 'Fat in Starbucks Drinks', 
     ylab = 'Total Sugar in Starbucks Food',
     main = 'Total Fat Vs Sugar for Starbucks Drinks', 
     data = drinks_expanded)
abline(h = 38, col = 'red')
```

### Plot StarBucks Drink (Sugars and Caffine)
```{r, warning=FALSE, message=FALSE}
# Blue line Max recomendation of sugar
# Red line Max recomendation of caffine

# https://www.fda.gov/consumers/consumer-updates/spilling-beans-how-much-caffeine-too-much

sudo<-drinks_expanded %>%
  mutate(clean_caff=as.numeric(as.character(drinks_expanded$Caffeine)))
sudo[is.na(sudo)]<-0

plot(clean_caff ~ Sugars, 
     ylab = 'Caffeine in Starbucks Drinks', 
     xlab = 'Sugar in Starbucks Food',
     main = 'Total Fat Vs Sugar for Starbucks Drinks', 
     data = sudo)
abline(h = 400, col = 'red')
abline(v = 38, col = 'blue')

#Venti Coffee
sudo %>%
  select(Beverage_category, clean_caff) %>%
  filter(clean_caff >= 400)

sudo %>%
  select(Beverage_category, Sugars) %>%
  filter(Sugars >= 38) %>% 
  count(Beverage_category) %>% 
  ungroup() %>% 
  arrange(desc(n))

sudo_sugar <- sudo %>%
  select(Beverage_category, Sugars) %>%
  filter(Sugars >= 38) 

# 37.60331% of drinks have higher than rated sugar
length(sudo_sugar[, 1]) / length(drinks_expanded$Beverage_category)

# Worst drinks ordered by violation of sugar contents
# Frappuccino Blended Coffee        
# Signature Espresso Drinks       
# Tazo Tea Drinks  
```

#### Checking for missing values
It doesnt seems the data contains any missing values. However, Caffeine variable in the `Drinks Expanded` dataset seems to be a mixed of numeric and character. As it can be seen, there are values with value `varies` or `Varies` in the data. Removed them from the dataset and will format it into numeric variable.

#### Distribution of the variables in each of the datasets
Drinks expanded dataset

```{r ggplot, fig.width=14, fig.height=8, warning=FALSE, message=FALSE, echo=FALSE}
drinks_expanded %>% 
    select(-contains("Beverage")) %>% 
    mutate(dummy = 1) %>% 
    melt(id = 'dummy') %>%
    mutate(value = as.numeric(value)) %>% 
    ggplot(aes(value)) + 
    geom_histogram(color = 'white', fill = 'red') + 
    facet_wrap(~variable, scales = 'free', ncol = 5) + 
    theme_bw() +
    labs(x = 'Value', y = 'Frequency', title = 'Drinks Expanded Variables Histograms')+
    theme(
        legend.text = element_text(size = 11),
        legend.position = "top",
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        strip.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        plot.title = element_text(size = 15))
```

###Food dataset
```{r fig.width = 14, fig.height = 4, warning = FALSE, message = FALSE, echo = TRUEALSE}
sbux_f %>% 
    select(-contains('Item')) %>%
    select(-contains('Type')) %>% 
    mutate(dummy = 1) %>% 
    melt(id = 'dummy') %>%
    mutate(value = as.numeric(value)) %>% 
    ggplot(aes(value)) + 
    geom_histogram(color = 'white',fill = 'red') + 
    facet_wrap(~variable, scales = 'free', ncol = 5) + 
    theme_bw() +
    labs(x = 'Value', y = 'Frequency', title = 'Food Variables Histograms') +
    theme(
        legend.text = element_text(size = 11),
        legend.position = "top",
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        strip.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        plot.title = element_text(size = 15))
```

### Supervised learning
Fit linear model to predict number of Calories in a drink based on other variables such as Total_Fat, Trans_Fat. Sugars etc.
5 folds cross validation Used and will use `MAPE` (mean average error percentage) as the perforamce metrics. Once we have the `MAPE` for each fold, I will simply take the average of all folds and this will be the final `MAPE` 
```{r warning = FALSE, message = FALSE}
# Randomly shuffle the data, set seed to produce the same resultss
# in addition, remove the first 3 columns which contains the Beverage descriptions
drinks_expanded <- drinks_expanded[sample(nrow(drinks_expanded)), ] %>% select(-contains("Beverage"))

# Create 5 equally size folds
K <- 5
folds <- cut(seq(1, nrow(drinks_expanded)), breaks = K, labels = FALSE)

# Perform 5 folds cross validation
MAPE <- c()
for(i in 1:K){
    index <- which(folds == i,arr.ind = TRUE)
    train <- drinks_expanded[-index, ]
    test <- drinks_expanded[index, ]
    model <- lm(Calories ~., data = train)
    prediction <- predict(model, test)
    MAPE[i] <- abs(test$Calories / prediction - 1)
}

paste0('MAPE for 5 folds: ', round(mean(MAPE) * 100, 2), '%')
```

### Unspurviesd learning
for the unspurviesd part, I will use the sbux_f dataset.
We will use "Hierarchical Clustering" in order to find clusters of similiar sbux_fs based on
`Calories`, `Fat`, `Carb`, `Fiver` and `Protein`
```{r fig.width = 24, fig.height = 15, warning = FALSE, message = FALSE}
# remove the sbux_f column 
df <- sbux_f %>% select(-Item, -Type)
# set sbux_f column as the row names
rownames(df) <- sbux_f$Item

# scale the data and create distance metrics using euclidean distance.
dd <- dist(scale(df), method = "euclidean")

# fit hierarchical clustering using "ward.D2" method
hc <- hclust(dd, method = "ward.D2")

# plot the results
par(cex = 1, mar = c(2, 2, 4, 1))
plot(hc, hang = -1, cex = 1.4, xlab = "", sub="", 
     axes=FALSE, ylab = '', main='Food HierarChial Cluster Diagram', cex.main = 2)
```

### Simulation
```{r}
# https://www.accessdata.fda.gov/scripts/interactivenutritionfactslabel/carbohydrate.html

starbucks_order <- function(food_size = 1, drink_size = 1){
  food_samp <- sample(sbux_f, food_size, replace = TRUE)
  drink_samp <- sample(drinks_expanded, drink_size, replace = TRUE)
  
  total_sugar <- sum(drink_samp$Sugars)
  total_caffine <- sum(drink_samp$Caffeine)
  total_calories <- sum(food_samp$Calories) + sum(drink_samp$Calories)
  total_fat <- sum(food_samp$Fat) + sum(drink_samp$Total_Fat)
  total_carbo <- sum(food_samp$Carb) + sum(drink_samp$Total_Carbohydrates)
  
  # Index 1 Sugar
  # Index 2 Calories
  # Index 3 Fat
  # Index 4 Carbohydrates
  # Index 5 Caffine
  return(c(total_sugar, total_calories, total_fat, total_carbo, total_caffine))
}

food_orders <- 3
drink_orders <- 3
n <- 365
final_sim <- do(n) * starbucks_order(food_orders, drink_orders)
head(final_sim)

hist(final_sim[, 1],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Sugar Intake at Starbucks',
     main = 'Daily Sugar Amount (g)')
abline(v = 38,col = 'red')

hist(final_sim[, 2],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Calories Intake at Starbucks',
     main = 'Daily Calories Amount')

hist(final_sim[, 3],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Fat Intake at Starbucks',
     main = 'Daily Fat Amount (g)')
abline(v = 65,col = 'red')

hist(final_sim[, 4],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Carbohydrates Intake at Starbucks',
     main = 'Daily Carbohydrates Amount (g)')
abline(v = 300,col = 'red')

hist(final_sim[, 5],
     breaks = 30,
     xlab = 'Graph of Simulated Daily Caffine Intake at Starbucks',
     main = 'Daily Caffine Amount (mg)')
abline(v = 400,col = 'red')
```