---
title: "STAT 380 Final Project"
subtitle: "Spring 2019"
author: "Sanchit Goel"
date: "Due: 04/30/2019 11:59PM"
output: html_notebook
---

# Front matter
```{r}
# always clean up R environment
rm(list = ls())

# load all packages here
library(dplyr)
library(tidyr)
library(ggplot2)
library(caret)
library(readr)
library(tidyverse)
library(ggthemes)
library(stringr)
library(repr)
library(plotly)
library(corrplot)
library(caTools)
library(rpart)
library(rpart.plot)

# Setting a random seed
set.seed(1)
```

```{r echo = F}
# User Defined Function:
replace_drop_na <- function(arg_1) {
   arg_1[arg_1 == '-'] <- NA
   print(paste0("Pre-Cleaning Records: ", nrow(arg_1)))

  # Dropping NA values records
  result <- arg_1 %>%
  drop_na()

  print(paste0("Cleaned Records: ", nrow(result)))
  return(result)
}
```


# Loading and Cleaning the Datasets
### McDonalds Dataset
```{r}
mcd <- read.csv(file = "nutrition_facts_mcdonalds_menu.csv", header = TRUE, sep = ",")
mcd <- replace_drop_na(mcd)

# Subsetting drinks from the menu
drinks.oz <- mcd[str_detect(mcd$Serving.Size, " fl oz.*"),]
drinks.ml <- mcd[str_detect(mcd$Serving.Size, 'carton'),]

# Converting ounces to mililiters and further rounding the values
drinks.oz$Serving.Size <- 
  round(as.numeric(gsub(" fl oz.*", "", drinks.oz$Serving.Size))*29.5735, 0)
drinks.ml$Serving.Size <- 
  round(as.numeric(gsub(".*\\((.*)\\ ml).*", "\\1", drinks.ml$Serving.Size)), 0)

# Keeping the values and removing characters
food.g <- mcd[str_detect(mcd$Serving.Size, 'g'),] 
food.g$Serving.Size <- 
  round(as.numeric(gsub(".*\\((.*)\\ g).*", "\\1", food.g$Serving.Size)), 0)

# Combining drinks and food menu
# While adding a new feature for drink or food
cleaned_mcd <- rbind(drinks.oz,drinks.ml)
cleaned_mcd$Type <- rep("Drinks", nrow(cleaned_mcd))
food.g$Type <- rep("Food", nrow(food.g))
cleaned_mcd <- rbind(cleaned_mcd,food.g)

keeps <- c("Item", "Calories", "Total.Fat", "Carbohydrates", "Dietary.Fiber", "Protein", "Type")
cleaned_mcd <- cleaned_mcd[keeps]
colnames(cleaned_mcd)[colnames(cleaned_mcd)=="Total.Fat"] <- "Total_Fat"
colnames(cleaned_mcd)[colnames(cleaned_mcd)=="Dietary.Fiber"] <- "Fiber"

cleaned_mcd$Calories <- as.numeric(as.character(cleaned_mcd$Calories))
cleaned_mcd$Total_Fat <- as.numeric(as.character(cleaned_mcd$Total_Fat))
cleaned_mcd$Carbohydrates <- as.numeric(as.character(cleaned_mcd$Carbohydrates))
cleaned_mcd$Fiber <- as.numeric(as.character(cleaned_mcd$Fiber))
cleaned_mcd$Protein <- as.numeric(as.character(cleaned_mcd$Protein))
```

### Starbucks Dataset
```{r}
# Food Dataset
sbux_f <- read.csv(file = "nutrition_facts_starbucks_food_menu.csv", header = TRUE, sep = ",")
sbux_f <- replace_drop_na(sbux_f)

colnames(sbux_f)[colnames(sbux_f)=="ï..Food.Name"] <- "Item"
colnames(sbux_f)[colnames(sbux_f)=="Fat..g."] <- "Total_Fat"
colnames(sbux_f)[colnames(sbux_f)=="Carb...g."] <- "Carbohydrates"
colnames(sbux_f)[colnames(sbux_f)=="Fiber..g."] <- "Fiber"
colnames(sbux_f)[colnames(sbux_f)=="Protein..g."] <- "Protein"

# Drink Dataset
sbux_d <- read.csv(file = "nutrition_facts_starbucks_drink_menu.csv", header = TRUE, sep = ",")
sbux_d <- replace_drop_na(sbux_d)

sbux_d <- sbux_d %>%
  select(-one_of(c('Sodium')))
colnames(sbux_d)[colnames(sbux_d)=="X"] <- "Item"
colnames(sbux_d)[colnames(sbux_d)=="Fat..g."] <- "Total_Fat"
colnames(sbux_d)[colnames(sbux_d)=="Carb...g."] <- "Carbohydrates"
colnames(sbux_d)[colnames(sbux_d)=="Fiber..g."] <- "Fiber"

# Adding a new column named Type
sbux_d$Type <- rep("Drinks", nrow(sbux_d))
sbux_f$Type <- rep("Food", nrow(sbux_f))
cleaned_sbux <- rbind(sbux_f, sbux_d)

cleaned_sbux$Calories <- as.numeric(as.character(cleaned_sbux$Calories))
cleaned_sbux$Total_Fat <- as.numeric(as.character(cleaned_sbux$Total_Fat))
cleaned_sbux$Carbohydrates <- as.numeric(as.character(cleaned_sbux$Carbohydrates))
cleaned_sbux$Fiber <- as.numeric(as.character(cleaned_sbux$Fiber))
cleaned_sbux$Protein <- as.numeric(as.character(cleaned_sbux$Protein))
```

### Joining the Mcdonalds and Starbucks Dataset
```{r}
# Joining the Food and Drink Menu for Starbucks and Mcdonalds
cleaned_sbux$Brand <- rep("Starbucks", nrow(cleaned_sbux))
cleaned_mcd$Brand <- rep("McDonalds", nrow(cleaned_mcd))
joined_data <- rbind(cleaned_mcd, cleaned_sbux)
```

### McDonalds Data Summary
```{r echo = F}
summary(cleaned_mcd)
```

### Starbucks Data Summary
```{r echo = F}
summary(cleaned_sbux)
```

### Joined Data Summary
```{r echo = F}
summary(joined_data)
```

```{r}
# Freeing up Memory of redundant data
rm('mcd', 'sbux_d', 'sbux_f', 'keeps', 'drinks.ml', 'drinks.oz', 'food.g')
```


## McDonalds:
#### Calories Histogram
```{r echo = F}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(x = Calories)) +
  geom_histogram(aes(y = ..density..), fill = "lightblue", binwidth = 40, color = "darkgray") + 
  scale_x_continuous(breaks = seq(min(cleaned_mcd$Calories), max(cleaned_mcd$Calories), by = 200)) + 
  geom_density()
```

#### Calories BoxPlot
```{r echo = F}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(x = reorder(Type, Calories), y = Calories)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = NULL)
```

#### Correlation Plot
```{r echo = F}
cr <- cor(cleaned_mcd[,2:6])
corrplot(cr, method = 'color')
rm('cr')
```

#### Calories Vs Total Fat Scatter Plot
```{r echo = F}
plot(cleaned_mcd$Total_Fat, cleaned_mcd$Calories, main = "Scatter Plot", xlab = "Total Fat", ylab = "Calories")
abline(lm(cleaned_mcd$Calories~cleaned_mcd$Total_Fat))
```

#### Colries vs Protein Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(y = Calories, x = Protein)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(y = Calories, x = Carbohydrates)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_mcd, aes(y = Calories, x = Total_Fat)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

### Model Training
#### Splitting the dataset
```{r}
split <- sample.split(cleaned_mcd$Calories, SplitRatio = 0.80)
train <- subset(cleaned_mcd, split == "TRUE")
test <- subset(cleaned_mcd, split == "FALSE")
```

#### Linear Regression Model
```{r}
mcd_lm_model <- lm(Calories ~ Total_Fat + Protein + Carbohydrates, data = train)
summary(mcd_lm_model)
```

#### Plots to explain the model
```{r echo = F}
plot(mcd_lm_model, which = 1:4)
```

#### Predictions vs Actual
```{r echo = F}
predictions <- predict(mcd_lm_model, test)
plot(test$Calories,type = "l",lty = 1.8, col = "blue")
lines(predictions,type = "l", col = "red")
```

#### Future Predictions
```{r}
predict(mcd_lm_model,data.frame(Total_Fat = 20, Protein = 18, Carbohydrates = 33))
```

#### Decision Tree
```{r}
mcd_dt_model <- rpart(Type ~ Calories + Total_Fat + Protein + Carbohydrates, data = train)
prp(mcd_dt_model, type = 1, extra = 3, main = "Decision Tree")

predictions <- data.frame(data = factor(predict(mcd_dt_model, test, type="class")), type = "prediction")
real <- data.frame(data = factor(test$Type), type = "real")
conf_data <- rbind(predictions, real)

# Check if the levels are identical
identical(levels(conf_data[conf_data$type == "prediction", 1]) , levels(conf_data[conf_data$type == "real", 1]))

confusionMatrix(conf_data[conf_data$type == "prediction", 1], conf_data[conf_data$type == "real",1], dnn = c("Prediction", "Reference"))
```

```{r}
# Freeing up Memory of redundant data
rm('predictions', 'real', 'conf_data', 'split', 'test', 'train')
```

## Starbucks:
#### Calories Histogram
```{r echo = F}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(x = Calories)) +
  geom_histogram(aes(y = ..density..), fill = "lightblue", binwidth = 40, color = "darkgray") + 
  scale_x_continuous(breaks = seq(min(cleaned_sbux$Calories), max(cleaned_sbux$Calories), by = 200)) + 
  geom_density()
```

#### Calories BoxPlot
```{r echo = F}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(x = reorder(Type, Calories), y = Calories)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = NULL)
```

#### Correlation Plot
```{r echo = F}
cr <- cor(cleaned_sbux[,2:6])
corrplot(cr, method = 'color')
rm('cr')
```

#### Calories Vs Total Fat Scatter Plot
```{r echo = F}
plot(cleaned_sbux$Total_Fat, cleaned_sbux$Calories, main = "Scatter Plot", xlab = "Total Fat", ylab = "Calories")
abline(lm(cleaned_sbux$Calories~cleaned_sbux$Total_Fat))
```

#### Colries vs Protein Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(y = Calories, x = Protein)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(y = Calories, x = Carbohydrates)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(cleaned_sbux, aes(y = Calories, x = Total_Fat)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Type, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

### Model Training
#### Splitting the dataset
```{r}
split <- sample.split(cleaned_sbux$Calories, SplitRatio = 0.80)
train <- subset(cleaned_sbux, split == "TRUE")
test <- subset(cleaned_sbux, split == "FALSE")
```

#### Linear Regression Model
```{r}
sbux_lm_model <- lm(Calories ~ Total_Fat + Protein + Carbohydrates, data = train)
summary(sbux_lm_model)
```

#### Plots to explain the model
```{r echo = F}
plot(sbux_lm_model, which = 1:4)
```

#### Predictions vs Actual
```{r echo = F}
predictions <- predict(sbux_lm_model, test)
plot(test$Calories,type = "l",lty = 1.8, col = "blue")
lines(predictions,type = "l", col = "red")
```

#### Future Predictions
```{r}
predict(sbux_lm_model,data.frame(Total_Fat = 20, Protein = 18, Carbohydrates = 33))
```

#### Decision Tree
```{r}
sbux_dt_model <- rpart(Type ~ Calories + Total_Fat + Protein + Carbohydrates, data = train)
prp(sbux_dt_model, type = 1, extra = 3, main = "Decision Tree")

predictions <- data.frame(data = factor(predict(sbux_dt_model, test, type="class")), type = "prediction")
real <- data.frame(data = factor(test$Type), type = "real")
conf_data <- rbind(predictions, real)

# Check if the levels are identical
identical(levels(conf_data[conf_data$type == "prediction", 1]) , levels(conf_data[conf_data$type == "real", 1]))

confusionMatrix(conf_data[conf_data$type == "prediction", 1], conf_data[conf_data$type == "real",1], dnn = c("Prediction", "Reference"))
```

```{r}
# Freeing up Memory of redundant data
rm('predictions', 'real', 'conf_data', 'split', 'test', 'train')
```

## Joined Data:
#### Calories Histogram
```{r echo = F}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(joined_data, aes(x = Calories)) +
  geom_histogram(aes(y = ..density..), fill = "lightblue", binwidth = 40, color = "darkgray") + 
  scale_x_continuous(breaks = seq(min(joined_data$Calories), max(joined_data$Calories), by = 200)) + 
  geom_density()
```

#### Calories BoxPlot
```{r echo = F}
options(repr.plot.height = 3, repr.plot.width = 6)
ggplot(joined_data, aes(x = reorder(Brand, Calories), y = Calories)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = NULL)
```

#### Correlation Plot
```{r echo = F}
cr <- cor(joined_data[,2:6])
corrplot(cr, method = 'color')
rm('cr')
```

#### Calories Vs Total Fat Scatter Plot
```{r echo = F}
plot(joined_data$Total_Fat, joined_data$Calories, main = "Scatter Plot", xlab = "Total Fat", ylab = "Calories")
abline(lm(joined_data$Calories~joined_data$Total_Fat))
```

#### Colries vs Protein Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(joined_data, aes(y = Calories, x = Protein)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Brand, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(joined_data, aes(y = Calories, x = Carbohydrates)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Brand, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

#### Colries vs Calories from Fat Density Map
```{r echo = F}
options(repr.plot.height = 4, repr.plot.width = 6)
ggplot(joined_data, aes(y = Calories, x = Total_Fat)) +
  geom_jitter(size = 1) +
  stat_density2d() +
  geom_smooth(method = 'lm', colour = 2) +
  facet_wrap( ~ Brand, scales = "free_x") +
  theme(panel.border = element_rect(colour = 1, fill = NA))
```

### Model Training
#### Splitting the dataset
```{r}
split <- sample.split(joined_data$Calories, SplitRatio = 0.80)
train <- subset(joined_data, split == "TRUE")
test <- subset(joined_data, split == "FALSE")
```

#### Linear Regression Model
```{r}
lm_model <- lm(Calories ~ Total_Fat + Protein + Carbohydrates, data = train)
summary(lm_model)
```

#### Plots to explain the model
```{r echo = F}
plot(lm_model, which = 1:4)
```

#### Predictions vs Actual
```{r echo = F}
predictions <- predict(lm_model, test)
plot(test$Calories,type = "l",lty = 1.8, col = "blue")
lines(predictions,type = "l", col = "red")
```

#### Future Predictions
```{r}
predict(lm_model,data.frame(Total_Fat = 20, Protein = 18, Carbohydrates = 33))
```

#### Decision Tree
```{r}
dt_model <- rpart(Type ~ Calories + Total_Fat + Protein + Carbohydrates, data = train)
prp(dt_model, type = 1, extra = 3, main = "Decision Tree")

predictions <- data.frame(data = factor(predict(dt_model, test, type="class")), type = "prediction")
real <- data.frame(data = factor(test$Type), type = "real")
conf_data <- rbind(predictions, real)

# Check if the levels are identical
identical(levels(conf_data[conf_data$type == "prediction", 1]) , levels(conf_data[conf_data$type == "real", 1]))

confusionMatrix(conf_data[conf_data$type == "prediction", 1], conf_data[conf_data$type == "real",1], dnn = c("Prediction", "Reference"))
```

```{r}
# Freeing up Memory of redundant data
rm('predictions', 'real', 'conf_data', 'split', 'test', 'train')
```